### 2025-11-04 paper/main.tex
weiter

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{caption}
\usepackage{csquotes}
\usepackage{float}
\usepackage{derivative}
\usepackage{subcaption}

\usepackage[
	style=nejm,
	citestyle=numeric-comp,
	sorting=none,
	doi=false,
	backend=bibtex]{biblatex}
\addbibresource{literature.bib}

\geometry{
	a4paper,
	total={6in, 9in},
	left=1.25in,
	top=1in,
}

\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}

\title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}
\author{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger\\
	\small University of Salzburg\\
	\small Department of Artificial Intelligence and Human Interfaces (AIHI) }

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
	\noindent
	We study how the numerical precision of weights, activations, and gradients in
	neural networks can be reduced without sacrificing too much of the performance.
	Using MNIST as a dataset, we benchmark uniform quantisation (mapping real
	values to evenly spaced levels via a single step size), mixed-precision
	formats, and post-training decimal rounding. We compare end-to-end
	low-precision training, inference-only quantisation, and mixed-precision with
	gradient clipping, and evaluate their impact on test accuracy and efficiency.
	The results on the used test setup show that (i) 16-bit end-to-end training
	does not significantly decrease test accuracy for the MNIST dataset, (ii)
	three-decimal-place weight rounding while keeping 32-bit precision incurs
	minimal accuracy loss. These findings verify limits for memory- and
	energy-constrained usecases like inference in mobile devices.
\end{abstract}

\section{Introduction}
Artificial Neural Networks (ANNs) are a key part of today's machine learning
systems. They have made huge progress in areas like image recognition, speech
processing, and self-driving cars. However, one of the biggest challenges with
large ANNs is that they use a lot of memory and computing power. This becomes a
problem when you want to run them on devices with limited resources, such as
smartphones, embedded systems, or edge devices. \\\\ In order to solve this,
researchers have looked into ways to make neural networks more efficient. A
common approach is to reduce the size of the network or shrink the input batch
size. But doing this can harm the network's ability to learn and generalize. A
better strategy is to reduce the numerical precision of the network's
parameters such as weights, activations, and gradients. This method, called
quantization, makes the model smaller and faster while using less energy,
without necessarily hurting accuracy. \\\\ An advanced version of this method
is called mixed-precision quantization. Instead of using the same number of
bits for every layer in the network, mixed-precision assigns different
bitwidths depending on how important each part is. For example, the early
layers that handle low-level features might use higher precision, while deeper
layers could work fine with fewer bits. This flexible approach often leads to
better results than using a single fixed precision for the whole model.
Researchers have used a range of techniques, like gradient descent, heuristics,
and even evolutionary algorithms, to find the best bitwidth settings for each
layer. \\\\ Another method that helps reduce resource usage is network pruning.
This means cutting out parts of the network that do not contribute much to its
performance. A way to do this is simulated annealing, a search algorithm that
mimics the process of cooling metal to find a stable structure. It does not
need gradient information and can still find good network shapes by exploring
many possibilities. This makes it useful when you want to shrink a network
without going through heavy retraining. \\\\ Previous studies have shown that
reducing the precision of neural network parameters can lead to significant
improvements in memory usage and processing speed. Even when precision is
reduced to very low levels, models can still perform competitively on complex
tasks. However, while these early results are promising, the overall effects of
lowering precision, especially how it influences model accuracy, training
behavior, and storage efficiency, are still not fully understood and deserve
more detailed investigation. \\\\ This is where our experimental work begins.
In this paper, we take a closer look at how different levels of parameter
precision affect the behavior and performance of neural networks. Our goal is
to understand what happens when we reduce precision in various ways, and how
much we can lower it before performance starts to drop. \\\\ We start by
testing how reducing precision during both training and inference affects model
accuracy and runtime. Next, we look at what happens when we train the network
using full precision, but lower the precision only during inference. This setup
is common in real-world applications, where a powerful machine handles
training, but the final model is deployed on a lightweight device. Then, we run
a series of tests where we round the trained weights to a limited number of
decimal places. This helps to see how much we can simplify the stored weight
values without hurting accuracy too much. This is especially useful for devices
with very limited storage. We also try out mixed-precision training, using a
mix of Half-Precision Floating Point (FP16) and Single-Precision Floating Point
(FP32) formats. On top of that, we experiment with different types of gradient
clipping, like global norm clipping and value clipping, to see how they affect
training stability and final accuracy. \\\\ Even though we use relatively
simple network architectures and datasets, our results show that neural
networks are quite robust to low-precision settings. In some cases, we were
able to round weights to just three decimal digits and still get nearly the
same accuracy as the full-precision version. These findings show that it is
possible to significantly reduce the memory and compute needs of neural
networks without losing much performance. This makes them more practical for
use in low-resource settings. It also opens the door for more advanced
techniques that combine quantization, pruning, and precision-aware training to
build highly efficient models for real-world applications.

\section{Artificial Neural Networks}

A typical neuron in a neural network performs a weighted sum of its inputs,
followed by an activation function (depicted in~\eqref{eq:activation}):

\begin{equation}\label{eq:activation}
	a = \omega^\top x + b,\quad \hat{y} = \sigma(a),
\end{equation}

where $\omega \in \mathbb{R}^n$ is the weight vector, $b \in \mathbb{R}$ is the
bias, $x \in \mathbb{R}^n$ is the input, and $\sigma$ is a nonlinear activation
function (e.g., sigmoid or ReLU). Training is conducted via backpropagation,
where gradients of a loss function $L(\hat{y}, y)$ with respect to the model
parameters are computed. For example, using the mean squared error:

\begin{align}\label{eq:mse}
	L(f(x_1, x_2)) & = {(\sigma(\omega^\top x + b) - y)}^2 \nonumber \\
	               & = {(\sigma(a) - y)}^2 \nonumber                 \\
	               & = {(\hat y - y)}^2
\end{align}
\begin{alignat}{3}\label{eq:grad-w1}
	\pdv{L}{\omega_1} & = \pdv{L}{\sigma} &  & \pdv{\sigma}{a}      &  & \pdv{a}{\omega_1} \nonumber \\
	                  & = 2(\hat y - y)~~ &  & \odv{\sigma(a)}{a}~~ &  & x_1
\end{alignat}

Memory and computational requirements during training are substantial, as
gradients, activations, and parameters must be stored and updated iteratively.
In contrast, inference requires only a forward pass, making it more amenable to
optimization via precision reduction.

\section{Memory Efficiency and Numerical Formats}

To reduce memory usage and computational load, precision-aware techniques such
as quantization and mixed-precision training are increasingly employed. Instead
of using 32-bit or 64-bit floating-point formats, parameters can be stored in
16-bit or even 8-bit fixed-point representations.
Figure~\ref{fig:floatingPointFormat} depicts the general floating point format
and terminology of significand, base, and exponent.
Tables~\ref{tab:binary-bits} and~\ref{tab:binary-precision} provide an overview
of the different floating point types and their respective characteristics.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{figures/float.png}
	\caption{Illustration of standard floating-point format.}\label{fig:floatingPointFormat}
\end{figure}

\begin{table}[H]
	\centering
	\begin{tabular}{@{}lcccc@{}}
		\toprule
		\textbf{Type} & Sign & Exponent & Significand & Total \\
		\midrule
		Half  (FP16)  & 1    & 5        & 10          & 16    \\
		Single (FP32) & 1    & 8        & 23          & 32    \\
		Double (FP64) & 1    & 11       & 52          & 64    \\
		\bottomrule
	\end{tabular}
	\caption{Field widths (in bits) of common IEEE754 formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-bits}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}{@{}lccc@{}}
		\toprule
		\textbf{Type} & Exponent bias & Precision (bits) & Decimal digits \\
		\midrule
		Half  (FP16)  & 15            & 11               & $\sim$3.3      \\
		Single (FP32) & 127           & 24               & $\sim$7.2      \\
		Double (FP64) & 1023          & 53               & $\sim$15.9     \\
		\bottomrule
	\end{tabular}
	\caption{Range and precision characteristics of the same formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-precision}
\end{table}

\section{Related Work}

Prior research in this domain can be broadly classified into two categories:

\subsection*{Quantization During Training}
Several studies have explored training neural networks with reduced precision
directly. Notable contributions include:
\begin{itemize}
	\item \textbf{Mixed Precision Training}, which leverages FP16 arithmetic for speed while maintaining FP32 master weights~\cite{micikevicius2017mixed}.
	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights~\cite{shen2020qbert}.
\end{itemize}

\subsection*{Post-Training Quantization}
Other approaches focus on converting pretrained models into lower-precision
equivalents:
\begin{itemize}
	\item \textbf{Integer-Only Quantization} using symmetric quantization scales~\cite{jacob2018quantization}.
	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining~\cite{frantar2022gptq}.
	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization~\cite{xiao2024smoothquant}.
\end{itemize}

\newpage

\section{Experimental Design}

We trained a simple fully connected neural network on the MNIST dataset,
consisting of 60,000 training and 10,000 test grayscale images of handwritten
digits (28×28 pixels, see figure~\ref{fig:mnistSample}). Model training was
performed under varying parameter bitwidths, and 5-fold cross-validation to
evaluate the model performance. The structure of the MLP is depicted in
figure~\ref{fig:mlpStructure}.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/mnist.png}
	\caption{Sample images from the MNIST dataset.}\label{fig:mnistSample}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/mlpStruc.png}
	\caption{MLP structure used in the experiments.}\label{fig:mlpStructure}
\end{figure}

\newpage
\section{Results}
In the first experiment, every floating-point tensor that is used inside the
training loop was converted to a lower precision data type. This therefore
includes the weights and biases, activations, gradients, optimizer states, etc.
The types tried were FP64, FP32, and FP16.

\subsection*{Accuracy Across Bitwidths}
Figure~\ref{fig:modelAccTrainAndInf} shows how the accuracy changed after the
models were trained on the different precision settings. Keep in mind that the
y-axis is not starting from zero, so the absolute height differences are
amplified visually, though the delta is only around $.1\%$. Re-training the
models and re-evaluating them has also lead to the accuracies being
``reversed'', meaning the FP64 model had the least accuracy. It is safe to
assume that the model accuracy has neither decreased, nor increased.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/trainingAndInference.png}
	\caption{Model accuracy evaluated after training with different precisions.}\label{fig:modelAccTrainAndInf}
\end{figure}

\subsection*{Training and Inference Time}
We also analyzed the times needed for both, training and inference.
Figure~\ref{fig:modelEvalTrainTimes} depicts the times measured in the training
loop and later on at the inference stage. Again, the times are almost identical
in both regards, and neither a real speed-up nor a slow-down has been measured.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/times.png}
	\caption{Training and evaluation times for the different precisions.}\label{fig:modelEvalTrainTimes}
\end{figure}

\subsection*{Inference-Only Testing}
This third experiment was conducted to test a possible post-training precision
reduction. For this, the model was trained on the highest precision (FP64) and
then quantized at the inference stage. Figure~\ref{fig:inferenceOnlyTesting}
shows that the accuracy of the baseline (FP64 training AND inference) is
maintained as the precision is decreased at the later stages. In contrast to
the \textit{Accuracy Across Bitwidths} experiment, there are no fluctuations at
all because for the lower precisions, there was no new model trained.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/inferenceOnly.png}
	\caption{Inference accuracy using FP64-trained weights followed by quantization.}\label{fig:inferenceOnlyTesting}
\end{figure}

\subsection*{Automatic Mixed Precision and Gradient Clipping}
Automatic Mixed Precision (AMP) accelerates deep learning training by using
both FP16 and FP32 data types reducing memory usage and increasing
computational speed. However, FP16's limited dynamic range can cause numerical
instability, especially during backpropagation. To address this, Gradient
Clipping is used to cap the magnitude of gradients (either globally or per
element) preventing extreme updates that could hinder convergence. Together,
AMP and gradient clipping enable efficient, stable training by balancing
performance and numerical precision.

First step is using global norm clipping, all gradients are collectively scaled
to ensure their combined norm does not exceed a specified threshold, promoting
stable updates across the network. In second step per-element clipping
constrains each individual gradient component within a fixed range, directly
limiting extreme values. These methods are evaluated under Automatic Mixed
Precision (AMP) to compare their impact on training stability, model accuracy,
and computation time across different precision types.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{figures/FP16_32_clipping.png}
	\caption{Affect of Gradient Norm on accuracy, training and testing timings.
		Accuracy is virtually the same between FP16 and FP32 when gradient clipping is used
		validating the stability of AMP and gradient scaling.
		Eval time is nearly identical for both formats, as evaluation involves fewer operations
		and less intense memory bandwidth usage.}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{figures/gradclip_16.png}
	\caption{Effect of Per-Element Gradient Clipping on FP16. Accuracy vs Training \& Eval Time}\label{fig:grad16}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{figures/gradclip_32.png}
	\caption{Effect of Per-Element Gradient Clipping on FP32. Accuracy vs Training \& Eval Time}\label{fig:grad32}
\end{figure}

Figure~\ref{fig:grad16} shows accuracy peaked at 97.07\% with a clipping value
of 0.5, while the lowest accuracy (96.55\%) occurred at clip=0.01, suggesting
that very small clip values may overly restrict gradient updates and hurt
learning. Training time ranged indicating that moderate clipping supports
efficient convergence. Evaluation time remained stable across all clip values
(around 2.01s to 2.16s), showing that clipping primarily impacts training
rather than inference performance. Also Figure~\ref{fig:grad32} shows that
accuracy was most stable across all clip values with the highest accuracy at
clip=0.5, demonstrating its robustness to gradient clipping variations.
Training time suggesting that larger clip values can speed up convergence
without sacrificing stability. Evaluation time confirming that clipping had
minimal effect on inference time.

\subsection*{Effect of Decimal Rounding}

A non training related issue is the storage of the NN parameters. Usually, the
parameters are stored in the format that was chosen to train with. This is best
for reproducibility as well as accuracy when doing inference with the trained
model weights. The datatype precision has a higher impact during training
though, which begs the question if one could omit precision from the model
weights. In this section, we explore a naive quantization method: Rounding to a
fixed number of decimals. With significantly little decimals and a reasonable
range for the integer part, one could store only the significant and provide
base as well as exponent as metadata. When loading the model, the weights are
converted back to floating point, with whatever precision is desired. Only the
weights are quantized; the input data, as well are the values propagated
through the network are at full precision.

The first step is to examine the distribution of parameter values. By the
standard deviation around modes one can make an educated guess as to how many
decimals are needed. The range will decide how large the integer part may get.
Reading from the distribution is just a heuristic though, so we will also test
the hypotheses.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/118_dist.png}
	\caption{
		Distribution of model weights in each layer. The values are in $[-0.5, 0.5]$,
		this means no integer part is needed in this example. Generally, with normalized data
		and the quite popular L1-/L2-/Batch-Norm layers, weight values in $[-1, 1]$ is to be expected.}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/118_dist2.png}
	\caption{Distribution of model weights in each layer after rounding to two decimals.
		This appears to be a good approximation already.}\label{fig:118_2}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/118_accuracy.png}
	\caption{Test Accuracy vs.\ number of decimal digits retained in weights.}\label{fig:118_acc}
\end{figure}

Figure~\ref{fig:118_acc} shows how rounding the decimal part affects the test
accuracy. For just one decimal the model has a maximum of 19 different weights
at its disposal, which makes the model have only 10\% accuracy, random guessing
with a 10 class problem. Two decimals perform significantly better. At 70\%
accuracy the model gets close to the full precision 88\% accuracy with 32 bit
floating point. Making and educated guess on distribution such as
Figure~\ref{fig:118_2} seems to be a good heuristic for this example. At just 3
decimals, the full 32 bit precision is reached. Thus we can conclude that with
just 10 bit for representing 3 decimals, no bits for the integer part and one
additional bit for the sign, only 11 bit must be stored to have the best
performing model for inference.

\section{Summary}
% \begin{itemize}
% 	\item \textbf{16-bit is enough for this classification task}: End-to-end FP16 preserved the accuracy.
% 	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
% \end{itemize}

% \section{Discussion}
% The results of the first experiments have shown that either the selected task (MNIST
% classification) was not complex enough to be affected by the decrease in accuracy, or the
% accuracy would need to be decreased even further than FP16. Otherwise, the results
% generally indicate that networks trained and evaluated with as low as 16-bit precision can
% achieve comparable performance to their high-precision counterparts. Moreover, rounding
% pretrained weights to just 3 decimal digits preserves full accuracy, opening the door to
% aggressive post-training quantization.This is particularly relevant for edge devices and
% real-time applications where memory and compute constraints are more relevant.

The experimental results presented in this paper demonstrate the robustness of
neural networks to reduced parameter precision, particularly for simpler tasks
like MNIST classification. Key findings include:

\begin{itemize}
	\item \textbf{Precision Reduction During Training and Inference}: Training and
	      evaluating models with 16-bit floating-point (FP16) precision preserved accuracy
	      comparable to higher-precision formats (FP32 and FP64). This suggests that for
	      tasks of similar complexity, FP16 is sufficient, offering potential memory and
	      computational savings without sacrificing performance.

	\item \textbf{Post-Training Quantization}: Reducing precision only during inference, after
	      training with high-precision (FP64), maintained baseline accuracy. This is
	      particularly valuable for deployment scenarios where training is done on powerful
	      hardware, but inference occurs on resource-constrained devices.

	\item \textbf{Mixed-Precision:} Training with mixed precision and Gradient Clipping
	      resulted in similar test accuracy scores as precision reduction during training
	      as well as post-training quantization. No significant decrease in model
	      performance accuracy was detected. The train and test times were similar,
	      regardless of using FP32, FP16 or a mix of both.

	\item \textbf{Weight Rounding:} Rounding trained weights to just three decimal digits
	      (approximately 11 bits) achieved full accuracy, highlighting the potential for
	      aggressive quantization in storage-efficient model deployment. This finding is
	      significant for edge applications where memory footprint is critical.
\end{itemize}

However, the simplicity of the MNIST dataset and the small network architecture
used in the experiments may limit the generalizability of these findings. More
complex tasks (e.g., CIFAR-10, ImageNet) and deeper architectures (e.g., CNNs,
Transformers) could exhibit greater sensitivity to precision reduction.
Additionally, the impact of precision on training dynamics, such as gradient
flow and convergence, was not deeply explored and could be a focus for future
work.

\section{Conclusion}

% This work confirms that for relatively simple tasks like MNIST classification,
% significant reductions in parameter precision can be achieved without
% compromising performance. Our future efforts will extend this analysis to more
% complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and
% explore advanced quantization-aware training techniques.

This study demonstrates that neural networks can maintain strong performance
even under significant reductions in parameter precision, particularly for
simpler tasks like MNIST classification. Key findings confirm that FP16
training and inference, post-training quantization, and aggressive weight
rounding (down to three decimal digits) preserve model accuracy while improving
memory and computational efficiency. These results are promising for deploying
models in resource-constrained environments, such as edge devices and real-time
applications.

To build on these insights and expand their applicability, future work should
focus on the following directions:

\begin{itemize}
	\item \textbf{Complex Architectures and Datasets:} Future research should validate these
	      findings on more complex models (e.g., CNNs, Transformers) and datasets (e.g., CIFAR-10,
	      ImageNet) to assess the scalability of low-precision techniques.

	\item \textbf{Quantization-Aware Training:} Investigating advanced methods like learned
	      quantization, non-uniform bit allocation, and dynamic precision adjustment could further
	      optimize the trade-off between efficiency and accuracy.

	\item \textbf{Hardware-Specific Optimizations:} Exploring how reduced precision interacts
	      with specialized hardware (e.g., TPUs, neuromorphic chips) could unlock new efficiency
	      gains in real-world deployments.

	\item \textbf{Theoretical Understanding:} A deeper analysis of how precision affects
	      gradient dynamics, convergence, and generalization could provide foundational insights for
	      designing robust low-precision training algorithms.
\end{itemize}

\newpage
\printbibliography%
\end{document}